---
title: "Analytics Project - Wellbeing Data Set"
subtitle: "C755: Analytics & Decision-Making in Healthcare"
author:
  name: Tirth Raval
  affiliation: C755 | DeGroote School of Business, McMaster University
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    number_sections: FALSE
    code_folding: hide
    toc: yes
    toc_float: 
      toc_collapsed: true
    theme: readable
    warning: false
---
# Load the Libraries 

```{r}
pacman::p_load(psych, xray , ggplot2, texreg, DT, wrapr, dplyr,
               sjmisc, sjlabelled, sjstats, sjPlot, tidyr,ggpubr,corrplot,
               knitr, kableExtra, captioner, car, excelR,ggcorrplot,Rmisc,sjmisc,funModeling,
               forcats, hrbrthemes,tidyverse,finalfit,patchwork,GGally,broom,caret,matrixTests,PerformanceAnalytics)



```  


# Connect the Dataset
```{r}
df = read.csv("Well_being_Health_Behavior_Environmental_Fact.csv")
```

```{r}
str(df)
```


```{r}
variables_recoded = c("wellbeing","SocialSupport","race","income","GeneralHealth","Asthma","HealthInsurance","CVD","LimitedActivity","Diabetes","employ","BMI","HeavyDrinker","CurrentSmoker","PhysicalActivity","marital")
for (i in variables_recoded){
  df[,i] <- as.factor(df[,i])
}
```


# Recoding the Variables
```{r}
df <- within(df, {
  wellbeing <- Recode(wellbeing, '"1" = "Poor_Wellbeing"; "0" = "Good_Wellbeing"', as.factor=TRUE, to.value="=", interval=":", 
  separator=";")
})
df <- within(df, {
  CurrentSmoker <- Recode(CurrentSmoker, '"1" = "Current Smoker"; "0" = "Not a Current Smoker"; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})
df <- within(df, {
  HeavyDrinker <- Recode(HeavyDrinker, '"1" = "Heavy Drinker"; "0" = "Not a Heavy Drinker"; ; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df <- within(df, {
  HealthInsurance <- Recode(HealthInsurance , '"1" = "With Health Coverage"; "0" = "No Health Coverage"', as.factor=TRUE, to.value="=", interval=":", 
  separator=";")
})
df <- within(df, {
  CVD              <- Recode(CVD, '"1" = "Cardiovascular Disease"; "0" = "No Cardiovascular Disease"; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})
df <- within(df, {
  Diabetes        <- Recode(Diabetes, '"1" = "Diabetic"; "0" = "Non Diabetic"; ; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df <- within(df, {
  LimitedActivity <- Recode(LimitedActivity , '"1" = "Limited"; "0" = "Not Limited"; ; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df <- within(df, {
  PhysicalActivity <- Recode(PhysicalActivity , '"1" = "Adequate Activity"; "0" = "Low Activity"; ; ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df <- within(df, {
  employ <- Recode(employ , '"1" = "Waged Employement"; "2" = "Self-Employed";"3" = "Unemployed for more than 1 yr" ;"4" = "Unemployed for less than 1 yr";"5" = "Homemaker";"6" = "Student";"7" = "Retired";"8" = "Unable to work" ;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df <- within(df, {
  BMI <- Recode(BMI , '"1" = "Low"; "2" = "Medium";"3" = "High" ;;;', as.factor=TRUE, to.value="=", 
  interval=":", separator=";")
})

df$income <- recode_factor(df$income, 
                                       "1" = "less than 15000", 
                                       "2" = "[15000-25000)", 
                                       "3" = "[25000-35000)",
                                       "4" = "[35000-50000)", 
                                       "5" = "50000 or more")

df$Asthma <- recode_factor(df$Asthma, '0' = "No", '1' = "Yes")

df$SocialSupport <- recode_factor(df$SocialSupport, '1' = "Never", '2' = " Rarely", '3' = " Sometimes", '4' = "Usually", '5' = "Always")

df$race <- recode_factor(df$race, '0' = "Other races", '1' = "White")

df$GeneralHealth <- recode_factor(df$GeneralHealth, '1' = "Excellent",
                                      '2' = "Very good",
                                      '3' = "Good",
                                      '4' = "Fair",
                                      '5' = "Poor")

```

# Interested Significant Predictors
```{r}
interested = c("wellbeing","SocialSupport", "GeneralHealth","PoorMentalHealthDays", "LimitedActivity", "employ")

df_interested = df[,interested]
```



# Interest and only omit values from significant predictors than the whole dataset
```{r}
df_interested <- na.omit(df_interested)
pacman::p_load(psych,DT)
headTail(df_interested, 10, 10) %>% datatable( rownames = FALSE, filter="top", options = list(pageLength = 21, scrollX=T))
cat(nrow(df_interested),ncol(df_interested))
```

# Logistic Regression Using Machine Learning

## Split the Data set

```{r}
## set the seed to make your partition reproducible
set.seed(123)
train_index <- sample(seq_len(nrow(df_interested)), size = 0.75*nrow(df_interested))

train <- df_interested[train_index, ]
test <- df_interested[-train_index, ]
nrow(train)

```

```{r}
nrow(test)
```


## Train the Model

```{r}
pacman::p_load(caret)
```

```{r}
glm_model <- train(wellbeing ~ SocialSupport+ GeneralHealth+PoorMentalHealthDays +LimitedActivity + employ, data = train, method = "glm", family = "binomial")
attributes(glm_model)
```


## Evaluate the Model Performance

```{r}
glm_model$metric
```


```{r}
glm_model$results
```


## Check the Predictions on the Test Data Set

```{r}
test$pred <- glm_model %>% predict(test)
```

## evaluate the performance-define accuracy function

```{r}
accuracy_function = function(real, predicted) {
  mean(real == predicted)
}
```

```{r}
accuracy_function(real = test$wellbeing,
         predicted = test$pred)
```

## Variable Importance of the Variables Used

```{r}
ggplot(varImp(glm_model))
```

## Performance Metrics for Log Reg ML Model
```{r}
confusionMtx_logreg = table(test$pred,test$wellbeing)
confusionMtx_logreg
```
### Create the Confusion Matrix for Log Reg

```{r}
TP_logreg=confusionMtx_logreg[1,1]
FP_logreg=confusionMtx_logreg[1,2]
TN_logreg=confusionMtx_logreg[2,2]
FN_logreg=confusionMtx_logreg[2,1]  
TP_logreg
```
### Calcaulating Recall

```{r}
recall_logreg=TP_logreg/(TP_logreg+FN_logreg)
recall_logreg
```
### Calcualting Accuracy

```{r}
Accuracy_logreg=(TP_logreg+TN_logreg)/nrow(test)
Accuracy_logreg
```
### Calacualting Specificity
```{r}
specificity_logreg=TN_logreg/(TN_logreg+FP_logreg)
specificity_logreg
```

### Calculate Precision
```{r}
precision_logreg=TP_logreg/(TP_logreg+FP_logreg)
precision_logreg
```
### Calculate F1-score
```{r}
F1_logreg=(2*recall_logreg*precision_logreg)/(recall_logreg+precision_logreg)
F1_logreg
```
### ROC for Log Reg ML Model   
```{r "retrieve probabilities of log reg ML model for ROC"}
# retrieve probabilities
prob_logreg<- glm_model %>% predict(test,type='prob')
```
```{r}
str(prob_logreg)
```

```{r "create logreg ML model roc"}
logreg.roc=roc(predictor=prob_logreg$Good_Wellbeing,response=test$wellbeing,
             levels=levels(test$wellbeing))
```

```{r}
logreg.roc
```

```{r "plot log reg ML model roc curve"}
ggroc(logreg.roc, alpha = 1, colour = "red", linetype = 1, size =0.7 ) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="black", linetype="dashed") + 
  ggtitle("ROC Curve for Logistic Regression ML Model") +
  xlab("1 - specificity") +
  ylab("sensitivity") +
  annotate(geom="text", x=0.7, y=0.75, label="Log Reg Model classifier", color="red") +  
  annotate(geom="text", x=0.45, y=0.4, label="random classifier") +
  annotate(geom="text", x=0.1, y=0.05, label="AUC = 0.9048", color="darkred") +
  theme_bw()
```
     
     

# CARET MODEL to predict Wellbeing


## Library Loading 
```{r}
library(rattle)
library(pROC)
library(xgboost)
library(randomForest)
```

## Split the Data
```{r}
set.seed(123)
train_index <- sample(seq_len(nrow(df_interested)), size = 0.75*nrow(df_interested))

train <- df_interested[train_index, ]
test <- df_interested[-train_index, ]
cat(nrow(test),nrow(train))
```



## Holdout

```{r}
levels(train$wellbeing) <- c("Good_Wellbeing", "Poor_Wellbeing")
```


```{r}
trcntrl_ho<-trainControl(classProbs = TRUE,summaryFunction = twoClassSummary)
```

## CART with Auto Hyperparameter Tuning




```{r}
model_CART <- train(wellbeing ~ SocialSupport+ GeneralHealth+PoorMentalHealthDays +LimitedActivity + employ,                 data = train, method = "rpart",trControl=trcntrl_ho,metric="ROC",tuneLength = 10)
model_CART$results
```


```{r}
plot(model_CART)
```



## Variable Importance of the CART Model

```{r}
ggplot(varImp(model_CART))
```


```{r}
print(varImp(model_CART))
```


## Plot the Tree

```{r fig.width=15}
plot(model_CART$finalModel, uniform=TRUE,
     main="Classification Tree")
text(model_CART$finalModel, all=FALSE, cex=.7)
```



```{r fig.width=15}
# Using the Rattle Library
library(rattle)
fancyRpartPlot(model_CART$finalModel,cex = 0.9)
```


## Prediction on Test Data

```{r}
test$pred_cart <- model_CART %>% predict(test)
```


```{r}
prob_cart<- model_CART %>% predict(test,type='prob')
```

```{r}
confusionMtx=table(test$pred_cart,test$wellbeing)
confusionMtx
```

## Create the Confusion Matrix

```{r}
TP=confusionMtx[1,1]
FP=confusionMtx[1,2]
TN=confusionMtx[2,2]
FN=confusionMtx[2,1]  
TP
```
### Calcaulating Recall

```{r}
recall=TP/(TP+FN)
recall
```
### Calcualting Accuracy

```{r}
Accuracy=(TP+TN)/nrow(test)
Accuracy
```
### Calacualting Specificity
```{r}
specificity=TN/(TN+FP)
specificity
```

### Calculate Precision
```{r}
precision=TP/(TP+FP)
precision
```
### Calculate F1-score
```{r}
F1=(2*recall*precision)/(recall+precision)
F1
```
```{r}
str(prob_cart)
```

```{r}
CART.roc=roc(predictor=prob_cart$Good_Wellbeing,response=test$wellbeing,
             levels=levels(test$wellbeing))
```

```{r}
CART.roc

```

```{r}
ggroc(CART.roc, alpha = 0.5, colour = "red", linetype = 2, size =0.7 )
```


```{r "plot CART ROC"}
ggroc(CART.roc, alpha = 1, colour = "red", linetype = 1, size =0.7 ) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="black", linetype="dashed") + 
  ggtitle("ROC Curve for CART Model") +
  xlab("1 - specificity") +
  ylab("sensitivity") +
  annotate(geom="text", x=0.7, y=0.75, label="CART Model classifier", color="red") +  
  annotate(geom="text", x=0.45, y=0.4, label="random classifier") +
  annotate(geom="text", x=0.1, y=0.05, label="AUC = 0.7415", color="darkred") +
  theme_bw()
```

